---
title: "Lab5"
author: "Luis J. Ramirez"
date: "February 25, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lab 5

**1. Loading in the credit data**
```{r, echo = TRUE}
credit <- read.csv("C:/Users/LuisJ/OneDrive/Desktop/RCode/MachineLearningR/data sets/credit.csv")
```

**2. Examining the structure of the data**

```{r,  echo = TRUE}
str(credit)
``` 
**3. Which features can be a possible predictor of a default loan ?**
It could be possible that employment duration, percent of income, checking balance, and months loan duration have a big impact on default loans

**4. Studying the following features**
```{r, echo = TRUE}
# pairs(credit$checking_balance, credit$employment_duration, credit$percent_of_income, credit$credit_history, credit$months_loan_duration, credit$default)
```
**5. Percentage of the loans that went into default**

```{r, echo = TRUE}
prop.table(table(credit$default))
```
**6.Why are prior lab approaches unwise**

**7. How to create a random sample in R** 
There are additional ways to create random samples from already stored data, an example is seen in **8.**
**8. Creating the random training and testing samples**


```{r, echo = TRUE}

set.seed(123)

train_sample <- sample(1000, 900)

credit_train <- credit[train_sample, ]
credit_test <- credit[-train_sample, ]

```

**9. Checking if the training train/test split s fair**
We check if the train/test split is fair to pervent training bias
```{r, echo = TRUE}
# train data
prop.table(table(credit_train$default))
#test data
prop.table(table(credit_test$default))
```

**10. Installing the c5.0 package**
This can be done in R-Studio enviroment or by using the following line in console
```{r, echo = TRUE}
#install.package(C50)
```

**11. Training the model **

```{r, echo = TRUE}
credit_model <- C50::C5.0(credit_train[-17], credit_train$default, trails = 1, costs = NULL)
credit_predition <- predict(credit_model, credit_test[-17])

summary(credit_model)
```


**12. We notice that if checking balance is less than 200 the loan is more likely to default, we also have ~15% error rate for type I and type II errors.**

**13. The tree is printed above.**

**14. - 16.** 
```{r, echo = TRUE}
#Evaluating model performance

library(caret)
library(e1071)
confusionMatrix(credit_predition, credit_test$default, dnn = c("Predicted", "Actual"))
```
**17. - 18. Improving the model using adaptive boosting**

```{r, echo = TRUE}
credit_boost10 <- C50::C5.0(credit_train[-17], credit_train$default, trails = 10, costs = NULL)
credit_boost10_pred <- predict(credit_boost10, credit_test[-17])

#summary(credit_boost10)

confusionMatrix(credit_boost10_pred, credit_test$default, dnn = c("Predicted", "Actual"))
```

**19. - 20. Cost matrix, we add a penalty for a type or error. When a the interest gained is better that the risk of giving out a loan, we actually want the loan. We notice that the accuracy decreased**
```{r, echo = TRUE}
matrix_dim <- list(c("no","yes"), c("no","yes"))
names(matrix_dim) <- c("predicted", "actual")
matrix_dim

# we suppose that a loan default cost the bank four times  as a missed oppurtunity

error_cost <- matrix (c(0,1,4,0), nrow = 2, dimnames = matrix_dim)

error_cost
```

```{r, echo = TRUE}
credit_boost10_err <- C50::C5.0(credit_train[-17], credit_train$default, trails = 10, costs = error_cost)
credit_boost10_err_pred <- predict(credit_boost10_err, credit_test[-17])

#summary(credit_boost10)

confusionMatrix(credit_boost10_err_pred, credit_test$default, dnn = c("Predicted", "Actual"))
```





